---
title: "R Notebook"
output: github_document
---


```{r message=FALSE}
library(jsonld)
library(jsonlite)
library(tidyverse)
library(printr)

```


```{r helper_fns, echo=FALSE}
is.url <- function(x){ 
  if(!is.character(x))
    return(FALSE)
  grepl("^http[s]?://.*", x)
}

add_context <- function(json, context = NULL){
  if(is.character(json)){ ## does not handle case of local files
    json <- jsonlite::fromJSON(json, simplifyVector = FALSE)
    if(length(json) > 1)
      json <- list("@graph" = json)
  }
  if(is.url(context)){
     json[["@context"]] <- context[["@context"]]
  } else { 
    if(is.character(context))
      context <- jsonlite::fromJSON(context, simplifyVector = FALSE)
    json[["@context"]] <- context[["@context"]]
  }
  jsonlite::toJSON(json, pretty=TRUE, auto_unbox = TRUE)
}

drop_context <- function(json){
  if(is.character(json)){
    json <- jsonlite::fromJSON(json, simplifyVector = FALSE)
  }
  json[["@context"]] <- NULL
  if(!is.null(json[["@graph"]]))
    json <- json[["@graph"]]
  jsonlite::toJSON(json, pretty=TRUE, auto_unbox = TRUE)

}

```

## Switching Contexts: Expansion & Compaction

- Good: homonyms (same word has different meanings), synonyms (different word has ~ same meaning)
- Fails: Mapping isn't 1:1

### A simple example

```{r}
ex <- 
 '{ 
      "date": "1993-10-02",
      "site": "N654",
      "scientificName": "Picea rubens",
      "area": 2,
      "count": 26
 }'
```

No need to alter your own data files/source data files.  Simply define a context that maps those terms into the reference vocabulary.  Data can use whatever fields it likes.


```{r}
map <- 
'{ 
  "@context": {
    "@vocab": "http://example.com/",
    "scientificName": "http://example.com/species"
  }
}'
target <- '{"@context": {"@vocab": "http://example.com/"}}'
add_context(ex, map) %>% 
  jsonld_compact(target)

```

### Using tabular schema

```{r}
df <- 
tribble(
  ~Date,       ~Site, ~Species, ~Area, ~Count,
  "10/1/1993", "N654", "PIRU",   2.0,     26,
  "10/3/1994", "N654", "PIRU",   2.0,     29,
  "10/1/1993", "N654", "BEPA",   1.0,     3
)

map <- 
'{ 
  "@context": {
    "@vocab": "http://example.com/",
    "Species": "http://example.com/scientificName"
  }
}'
target <- '{"@context": {"@vocab": "http://example.com/"}}'

df %>% 
  toJSON() %>% 
  add_context(map) %>% 
  jsonld_compact(target) %>% 
  drop_context() %>%
  fromJSON()

```
Wow does that seem like a round-about way to just do:

```{r}
df %>% rename(scientificName = Species)
```

and it is.  But a few things are worth noting:

- Robust to what columns are actually present.  Instead of a default vocabulary we can map all terms explicitly.  
  - Terms not expanded by the map (that is, not part of our context) will be dropped. (`Count` in the example below)
  - Terms not in our target context remain as external keys, and are not compacted.  
- Typesafe transforms. Terms (keys/column names) with the wrong type will not compact if the types do not match the target transform.  This way they are not actually lost, but cannot be confused.  

```{r}
map <- 
'{ 
  "@context": {
    "ex": "http://example.com/",
    "Species": "ex:scientificName",
    "Site": "ex:Site",
    "Date": "ex:Date",
    "Area": "ex:Area"
  }
}'

target <- '{
"@context": {
  "ex": "http://example.com/",
  "scientificName": "ex:scientificName",
  "Site": "ex:Site",
  "Count": "ex:count",
  "Date": {"@id": "ex:Date", "@type": "ex:Date"}
}}'

df %>% 
  toJSON() %>% 
  add_context(map) %>% 
  jsonld_compact(target) %>% 
  drop_context() %>%
  fromJSON() 


```


```{r}
map <- '{ "@context": { "@vocab": "https://data.berkeley.edu/" }}'

target <- '{
"@context": {
  "ucb": "https://data.berkeley.edu/",
  "hav": "http://data.harvard.edu/",
  "scientificName": "ucb:Species",
  "Site": "ucb:Site",
  "Count": "ucb:Count",
  "Date": "ucb:Date",
  "Area": "ucb:Area",
  "Density": "hav:density"
}}'

df %>% 
  toJSON() %>% 
  add_context(map) %>% 
  jsonld_compact(target) %>% 
  drop_context() %>%
  fromJSON() -> ucsb_df

ucsb_df %>% mutate(density = Count / Area)

```

----------




## Integration: Flattening shares id

```{r}
context <- '{   "@context": { "@vocab": "http://example.com/" } }'

ex <- '{
  "@context": { "@vocab": "http://example.com/" },
  "datasets": [
    {
      "@id": "http://global_unique_id",
      "@type": "dataset",
      "image": "http://image_uri",
      "name": "stuff"
    },
  
    {
      "@type": "dataset",
      "@id": "http://global_unique_id",
      "name": "Species name"
    }
  ]
}'

jsonld_flatten(ex, context)

```


--------

hdf5: column data, with units, metadata
