---
output: github_document
---

[![stability-experimental](https://img.shields.io/badge/stability-experimental-orange.svg)](https://github.com/joethorley/stability-badges#experimental)
[![Travis-CI Build Status](https://travis-ci.org/cboettig/emld.svg?branch=master)](https://travis-ci.org/cboettig/emld) 
[![Coverage Status](https://img.shields.io/codecov/c/github/cboettig/emld/master.svg)](https://codecov.io/github/cboettig/emld?branch=master)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# emld

The goal of emld is to provide a way to work with EML metadata in the JSON-LD format. At it's heart, the package is simply a way to translate an EML XML document into JSON-LD and be able to reverse this so that any semantically equivalent JSON-LD file can be serialized into EML-schema valid XML.  The package has only three core functions:

- `as_emld()` Convert EML from `xml` (or `json`) into a native R object, `emld` (list / S3 class).  
- `as_xml()` Convert the native R format, `emld`, back into XML-schema valid EML.
- `as_json()` Convert the native R format, `emld`, into `json`(LD).


This is very much **work in progress** The outline below illustrates things we can do or will be able to do with this package.  The examples below are just a sketch of ideas so far, I hope to replace these with richer examples that will probably be developed more fully as vignettes.  
## Installation

You can install emld from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("cboettig/emld")
```



## Motivation


In contrast to the existing [EML package](https://ropensci.github.io/EML), this package aims to a very light-weight implementation that seeks to provide both an intuitive data format and make maximum use of existing technology to work with that format.  In particular, this package emphasizes tools for working with linked data through the JSON-LD format.

Note that the JSON-LD format is considerably less rigid than the EML schema.  This means that there are many valid, semantically equivalent representations on the JSON-LD side that must all map into the same or nearly the same XML format.  At the extreme end, the JSON-LD format can be serialized into RDF, where everything is flat set of triples (e.g. essentially a tabular representation), which we can query directly with semantic tools like SPARQL, and also automatically coerce back into the rigid nesting and ordering structure required by EML.  This ability to "flatten" EML files can be particularly convenient for applications consuming and parsing large numbers of EML files.  This package may also make it easier for other developers to build on the EML, since the S3/list and JSON formats used here have proven more appealing to many R developers than S4 and XML serializations.

JSON-LD also makes it easier to extend EML with other existing semantic vocabularies.  The standard JSON-LD operations (e.g. framing, compaction) make it easy for developers to specify desired data structures, filter unnecessary terms and provide defaults for needed ones, or even define custom property names, rather than working with the often cumbersome prefixes or URIs of linked data. 



```{r}
library(emld)
library(jsonlite)
library(magrittr)
```


## Reading EML

The `EML` package can get particularly cumbersome when it comes to extracting and manipulating existing metadata in highly nested EML files.  The `emld` approach can leverage a rich array of tools for reading, extracting, and manipulating existing EML files. 

### Parse & serialize

We can parse a simple example and manipulate is as a familiar list object (S3 object):

```{r}
f <- system.file("extdata/example.xml", package="emld")
eml <- as_emld(f)
eml$dataset$title
```

```{r}
eml$dataset$title <- "A new title"
as_xml(eml, "test.xml")
```

We can prove that writing the list back into XML still creates a valid EML file. 

```{r}
eml_validate("test.xml")
```

 

### Query


We can query it with SPARQL, a rich, semantic way to extract data from one or many EML files.

```{r}
library(rdflib)
```

FIXME replace with an example(s) that makes better use of semantic relationships.

```{r}
f <- system.file("extdata/hf205.xml", package="emld")

as_emld(f) %>%
as_json("hf205.json")

sparql <- 
  'PREFIX eml: <http://ecoinformatics.org/>

  SELECT ?genus ?species ?northLat ?southLat ?eastLong ?westLong 

  WHERE { 
    ?y eml:taxonRankName "genus" .
    ?y eml:taxonRankValue ?genus .
    ?y eml:taxonomicClassification ?s .
    ?s eml:taxonRankName "species" .
    ?s eml:taxonRankValue ?species .
    ?x eml:northBoundingCoordinate ?northLat .
    ?x eml:southBoundingCoordinate ?southLat .
    ?x eml:eastBoundingCoordinate ?eastLong .
    ?x eml:westBoundingCoordinate ?westLong .
  }
'
  
rdf <- rdf_parse("hf205.json", "jsonld")
df <- rdf_query(rdf, sparql)
df
```

We can query it with JQ, a [simple and powerful query language](https://stedolan.github.io/jq/manual/) that also gives us a lot of flexibility over the return structure of our results:


```{r}
library(jqr)

as_emld(f) %>%
  as_json() %>% 
  as.character() %>%
  jq('.dataset.coverage.geographicCoverage.boundingCoordinates | 
       { northLat: .northBoundingCoordinate, 
         southLat: .southBoundingCoordinate }')
```

Some prototype examples of how we can use this to translate between EML and <http://schema.org/Dataset> representations of the same metadata can be found in <https://github.com/cboettig/emld/blob/master/notebook/jq_maps.md>


### Flatten and Compact.

We can flatten it, so we don't have to do quite so much sub-setting.  When we're done editing, we can compact it back into valid EML. 

```{r}
library(jsonld)
flat <- as_emld(f) %>%
  as_json() %>% 
  jsonld_flatten('{"@vocab": "http://ecoinformatics.org/"}') %>%
  fromJSON(simplifyVector = FALSE)
flat <- flat[["@graph"]]

```

FIXME this would be way more useful if nodes all had `@type` and were named by that type. Then we could do `flat$boundingCoordinates`. Currently flattened objects are unnamed and untyped, so this is less useful.  

## Writing EML

This section is even more experimental currently, and may not be a good direction for development.  Nevertheless, it can illustrate some of the convenience (and risk) of a simple S3 class.

Remember that `emld` objects are just nested lists, so you can create EML just by writing lists: 

```{r}

me <- list(individualName = list(givenName = "Carl", surName = "Boettiger"))

eml <- list(dataset = list(
              title = "dataset title",
              contact = me,
              creator = me),
            "#system" = "doi",
            "#packageId" = "10.xxx")

as_xml(eml, "ex.xml")
testthat::expect_true(eml_validate("ex.xml") )
```

Note that we don't have to worry about the order of the elements here, `as_xml` will re-order if necessary to validate. (For instance, in valid EML the `creator` becomes listed before `contact`.)  

Clearly most users are unlikely to remember what the possible properties are for each element, or which ones are attributes (which currently need the `#` prefix).  The `template()` function offers a quick way to address this issue:

```{r}
template("creator")
```


returns a simple list object, cast as an `emld` S3 class so that it prints cleanly.  Note that properties which take other objects are indicated by `{}` while properties taking text values are indicated by `~`.  If we assignt it to a variable instead of merely printing the output of `template()`, we can then fill out values as needed (making use of tab completion):

```{r}
scott <- template("creator")
scott$individualName <- template("individualName")
scott$individualName$givenName <- list("Scott", "A")
scott$individualName$surName <- "Chamberlain"
scott$electronicMailAddress <- "scott@ropensci.org"
```



We can add this element to the EML list we already created:

```{r}
eml$dataset$creator <- list(me, scott)
```

and confirm the file validates with this additional creator:

```{r}
as_xml(eml, "ex.xml")
testthat::expect_true(eml_validate("ex.xml") )
```


It remains to be seen if this proves substantially easier than EML creation with the S4-based constructors.  Some things are certainly more convienent: no `new`, no `ListOf` types, no remembering to index repeated elements, no `.Data` slots, etc.  On the other hand, S4 system prevents assigning to a slot an object that is of the wrong type.  The list approach does not have this safeguard. 

Ultimately it might be more useful as a developer tool to construct helper functions that can streamline creation of EML. 




```{r include=FALSE}
unlink("eml.xml")
unlink("test.xml")
unlink("ex.xml")
unlink("hf205.json")
```

### Going deeper

Several basic steps still to do: in particular, haven't yet expanded out `references` or EML's semantic annotations into more native JSON-LD (and vice-versa back into EML).  The latter in particular might be a convenient way to serialize additional semantic annotations onto an EML description. 





